ZCB Pricing

### 2018-08-01 ZCB Pricing .Rmd
Adjustments

# Crude Monte Carlo

@@

Crude_Monte_Carlo()
bond_price()
```

@@

Antithetic_variates = function(alpha =0.08,beta=0.8,sigma=0.005,T=5,r = 0.08,n=1000,s=100){
  dt <- T/s # Length of sub-intervals
  sim_path= matrix(0,s+1,n) # Matrix that stores rate pathways
  sim_path_inv= matrix(0,nrow = s+1, ncol = n)# matrix to generate interest paths using negative values of RV
  sim_path[1,] = r
  sim_path_inv[1,] = r
  z = rnorm(n) # Generate standard normal values
  z. = -z # Negative RV

  for (i in 1:n){
    for (j in 2:(s+1)){
      dr=(alpha-beta*sim_path[j-1,i])*dt+sigma*sqrt(dt)*z[i] # Same Euler Discretization as in the Crude Monte Carlo function
      dr_i= (alpha-beta*sim_path_inv[j-1,i])*dt + sigma*sqrt(dt)*z.[i] # Euler Discretization for negative rates
      sim_path[j,i]=sim_path[j-1,i]+dr
      sim_path_inv[j,i]=sim_path_inv[j-1,i]+dr_i

@@

  # Monte Carlo Simulations
  Sum=colSums(sim_path[2:(s+1),]*dt) # The Sum and Sum. variables record the summation of their respective rates to approximate the integral
  Sum.=colSums(sim_path_inv[2:(s+1),]*dt)
  int=exp(-Sum.) #Exponential of the estimate
  int.=exp(-Sum) #Exponential of the estimate under the negative rv 
  mean_1=mean(int) # Estimate using Z_i
  mean_2=mean(int.)  # Estimate using -Z_i
  result= (mean_1+mean_2)/2 # Final estimate
  result

@@

```{r 4}
#################################################################################
# Moment matching
#################################################################################

Moment_Matching=function(alpha =0.08,beta=0.8,sigma=0.005,T=5,r0 = 0.08,n=1000,s=100){

@@

```{r 5}
#################################################################################
# Control Variates
#################################################################################

@@

bond_price()
Control_Variates_estimate()
```

```{r}
# Comparisons of different methods
#################################################################################
# Stability of results
#################################################################################

CR=NULL
AN=NULL
MM=NULL
CV=NULL

@@

```{r 6}

#======================================================================================
# Doing some plots

xvec <- seq(-0.3,0.5,0.01)
S <- NULL
for (r in 1:length(xvec)){
  S[r] <- bond_price(0.08,0.8,0.005,0,5,xvec[r])
}

CR <- NULL
AV <- NULL
MM <- NULL
CO <- NULL

for (k in 1: length(xvec)){
  CR[k] <- Crude_Monte_Carlo(r0 = xvec[k], T=5)
}

for (s in 1: length(xvec)){
  AV[s] <- Antithetic_variates(r0 = xvec[s], T =5)

for (j in 1: length(xvec)){
  MM[j] <- Moment_Matching(r0 = xvec[j], T =5)

for (t in 1: length(xvec)){
  CO[t] <- Control_Variates_estimate(r0 = xvec[t], T =5)

plot(CR, x = xvec[1:length(xvec)],xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Monte Carlo vs Known Solution")
lines(xvec,S, col ="blue", lty =1.5)
lines(xvec,AV, col ="black", lty =1.5)
lines(xvec,MM, col ="yellow", lty =1.5)
lines(xvec,MM, col ="green", lty =1.5)
```

```{r 7}
# Explicit Method

explicit = function(alpha= 0.08,beta=0.8,sigma=0.005,r=0.08,I=1:1250, J =seq(-0.4,0.6,0.005),T=5){

  M <- length(J)
  N <- length(I)

  dt <- T/N
  dr <- J[2] -J[1]

@@

```{r 12, echo = FALSE}
# Comparison of results
# Varying short rates

compare_mat <- matrix(NA,nrow = 10, ncol = 10) # matrix of comparison

colnames(compare_mat) <- c("Alpha","Beta","Sigma","Initial rate","Known Price","Monte Carlo"," MC System Time","MC Squared Deviations", "Moment Matching", "System Time")

for (j in 1:10){
  compare_mat[j,1] <- 0.08
  compare_mat[j,2] <- 0.8
  compare_mat[j,3] <- 0.005
  compare_mat[j,4] <- 0.01 +0.01*(j-1)
  compare_mat[j,5] <- bond_price(alpha = compare_mat[j,1],beta = compare_mat[j,2], sigma = compare_mat[j,3], t= 0, T = 1, r = compare_mat[j,4])
  start_time <- proc.time()[1]
  compare_mat[j,6] <- Crude_Monte_Carlo(alpha = compare_mat[j,1],beta = compare_mat[j,2], sigma = compare_mat[j,3], T = 1, r0 = compare_mat[j,4], n = 1000, s=250)
  compare_mat[j,7] <- proc.time()[1] -start_time
  compare_mat[j,8] <- (compare_mat[j,6] -compare_mat[j,5])^2
  start_time <- proc.time()[1]
  compare_mat[j,9] <- Moment_Matching(alpha = compare_mat[j,1],beta = compare_mat[j,2], sigma = compare_mat[j,3], T = 1, r0 = compare_mat[j,4], n = 1000, s=250)
  compare_mat[j,10] <- proc.time()[1] - start_time

@@

```{r}
# Sensitivity Tests

@@

```{r}
#Stability
stables <- matrix(nrow=4, ncol=100) # matrix to hold prices for various parameters

for (i in 1:100){
  stables[1,i] <- Crude_Monte_Carlo()
  stables[2,i] <- Antithetic_variates()
}

for (i in 1:100){
  stables[3,i] <- Moment_Matching()
  stables[4,i] <- Control_Variates_estimate()
}

plot(1:100, stables[1,], type="l", col="blue")
lines(x=1:100, y=stables[2,], col="red")

>>>>

# Crude Monte Carlo function

@@

@@

Antithetic_variates = function(alpha =0.08,beta=0.8,sigma=0.005,T=5,r = 0.08,n=1000,s=100){

  dt <- T/s # Length of sub-intervals
  sim_path= matrix(0,s+1,n) # Matrix that stores rate pathways
  sim_path_inv= matrix(0,nrow = s+1, ncol = n)# matrix to generate interest paths using negative values of RV
  sim_path[1,] = r
  sim_path_inv[1,] = r
  z = rnorm(n) # Generate standard normal values
  z. = -z # Negative RV

  for (i in 1:n){
    for (j in 2:(s+1)){
      dr = (alpha-beta*sim_path[j-1,i])*dt+sigma*sqrt(dt)*z[i] # Same Euler Discretization as in the Crude Monte Carlo function
      dr_i = (alpha-beta*sim_path_inv[j-1,i])*dt + sigma*sqrt(dt)*z.[i] # Euler Discretization for negative rates
      sim_path[j,i] =s im_path[j-1,i]+dr
      sim_path_inv[j,i] = sim_path_inv[j-1,i]+dr_i

@@

  # Monte Carlo Simulations
  Sum = colSums(sim_path[2:(s+1),]*dt) # The Sum and Sum. variables record the summation of their respective rates to approximate the integral
  Sum.= colSums(sim_path_inv[2:(s+1),]*dt)
  int = exp(-Sum.) #Exponential of the estimate
  int.= exp(-Sum) #Exponential of the estimate under the negative rv 
  mean_1 = mean(int) # Estimate using Z_i
  mean_2 = mean(int.)  # Estimate using -Z_i
  result = (mean_1+mean_2)/2 # Final estimate
  result

@@

```{r 4}

# Moment matching

Moment_Matching = function(alpha = 0.08, beta = 0.8, sigma = 0.005, T = 5,r = 0.08, n = 1000, s = 200){

@@

```{r 5}

# Control Variates

@@

@@

```{r 6}
# Explicit Method

explicit = function(alpha = 0.08,beta = 0.8, sigma = 0.005,r = 0.08, I = 1:1250, J = seq(-0.4,0.6,0.005),T = 5){

  # I = partition of time interval
  # J = partition of short rates sequence

  M <- length(J) # Number of short rate grid points
  N <- length(I) # Number of time grid poiints

  dt <- T/N # Lenght of each time subinterval
  dr <- J[2] -J[1] # Lenght of each time interval

@@

```{r 9}
# Plots of the different methods
par(mfrow = c(2,2))

xvec <- seq(-0.4,0.6,0.005) # Vector for rates

known_sol <- NULL # Vector to hold the known closed form solution
for (r in 1:length(xvec)){
  known_sol[r] <- bond_price(0.08,0.8,0.005,0,5,xvec[r])
}

CR <- NULL # Vector to hold Crude Monte Carlo Solutions
AV <- NULL # Vector to hold Antithetic Monte Carlo Solutions
MM <- NULL # Vector to hold Moment Matching  Carlo Solutions
CV <- NULL # Vector to hold Control Variates Monte Carlo Solutions

Expl <- NULL # Vector to hold Explicit method Solutions
Impl <- NULL # Vector to hold Implicit method Solutions
Crank <- NULL # Vector to hold Crank-Nicolson Solutions

# Plots of Monte Carlo Methods ----------------------------------------------------------------------

for (i in 1: length(xvec)){
  CR[i] <- Crude_Monte_Carlo(r = xvec[i])
}

for (i in 1: length(xvec)){
  AV[i] <- Antithetic_variates(r = xvec[i])
}

for (i in 1: length(xvec)){
  MM[i] <- Moment_Matching(r = xvec[i])
}

for (i in 1: length(xvec)){
  CV[i] <- Control_Variates_estimate(r = xvec[i])
}

plot(CR, x = xvec[1:length(xvec)],xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Monte Carlo Solutions vs Known Solution")
lines(xvec, known_sol, col ="blue", lty =1.5)
lines(xvec, AV, col ="black", lty =1.5)
lines(xvec, MM, col ="yellow", lty =1.5)
lines(xvec, CV, col ="green", lty =1.5)

# Plots of Finite Difference Methods ----------------------------------------------------------------------

for (i in 1: length(xvec)){
  Expl[i] <- explicit(r = xvec[i])
}

for (i in 1: length(xvec)){
  Impl[i] <- implicit(r = xvec[i])
}

for (i in 1: length(xvec)){
  Crank[i] <- CN(r = xvec[i])
}

plot(Expl, x = xvec ,xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Explicit Method vs Known Solution")
lines(xvec, known_sol, col ="blue", lty =1.5)

plot(Impl, x = xvec ,xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Implicit Method vs Known Solution")
lines(xvec, known_sol, col ="blue", lty =1.5)

plot(Crank, x = xvec ,xlab = " Annual Interest Rate",ylab = "Bond Price",type = "l", col ="red", main = "Crank-Nicolson Method vs Known Solution")
lines(xvec, known_sol, col ="blue", lty =1.5)

```{r 10, echo = FALSE}
# Comparison of results for numerical methods

compare_mat <- matrix(NA,nrow = 7, ncol = 4) # matrix of comparison

colnames(compare_mat) <- c("Price", "Known Price", "Error", "System Time")
rownames(compare_mat) <- c("Crude Monte Carlo", "Antithetic Variates", "Moment Matching", "Control Variates", "Explicit Method", "Implicit Method", "Crank-Nicolson Method")

compare_mat[,2] <- bond_price()

# Crude Monte Carlo
start_time <- proc.time()[1]
compare_mat[1,1] <- Crude_Monte_Carlo()
compare_mat[1,4] <- proc.time()[1] - start_time
compare_mat[1,3] <- compare_mat[1,2] - compare_mat[1,1]

# Antithetic Variates
start_time <- proc.time()[1]
compare_mat[2,1] <- Antithetic_variates()
compare_mat[2,4] <- proc.time()[1] - start_time
compare_mat[2,3] <- compare_mat[2,2] - compare_mat[2,1]

# Moment Matching
start_time <- proc.time()[1]
compare_mat[3,1] <- Moment_Matching()
compare_mat[3,4] <- proc.time()[1] - start_time
compare_mat[3,3] <- compare_mat[3,2] - compare_mat[3,1]

# Control Variates
start_time <- proc.time()[1]
compare_mat[4,1] <- Control_Variates_estimate()
compare_mat[4,4] <- proc.time()[1] - start_time
compare_mat[4,3] <- compare_mat[4,2] - compare_mat[4,1]

# Explicit method
start_time <- proc.time()[1]
compare_mat[5,1] <- explicit()
compare_mat[5,4] <- proc.time()[1] - start_time
compare_mat[5,3] <- compare_mat[5,2] - compare_mat[5,1]

# Implicit method
start_time <- proc.time()[1]
compare_mat[6,1] <- implicit()
compare_mat[6,4] <- proc.time()[1] - start_time
compare_mat[6,3] <- compare_mat[6,2] - compare_mat[6,1]

# Crank-Nicolson method
start_time <- proc.time()[1]
compare_mat[7,1] <- CN()
compare_mat[7,4] <- proc.time()[1] - start_time
compare_mat[7,3] <- compare_mat[7,2] - compare_mat[7,1]

@@

```{r 11}
# Sensitivity Tests

@@

```{r 12}
#Stability

CR = NULL
AN = NULL
MM = NULL
CV = NULL
